Left-Linear Rewriting in Adhesive Categories

### 2024-10-16 concur/presentazione/pres_lunga.tex
finale

\newcommand{\modd}[0]{\catname{Mod}}
\newcommand{\des}[1]{Des({#1})}
\newcommand{\prof}[2]{\mathsf{Proof}({#1},{#2})}
\newcommand{\ab}[0]{=-Adj}
\newcommand{\ba}[0]{=-E}
\newcommand{\su}[2]{\mathfrak{s}_{#1}^{#2}}
\newcommand{\ipos}[1]{\catname{Pos}(\catname{#1})}
\newcommand{\Su}[2]{\mathfrak{S}_{#1}^{#2}}
\newcommand{\Pro}[2]{\mathfrak{P}_{#1}^{#2}}
\newcommand{\surr}[0]{\mathsf{surr}}
\newcommand{\propag}[0]{\mathsf{propag}}
\newcommand{\mdl}[1]{\catname{Mod}(#1)}
\newcommand{\alg}[1]{\catname{FAlg}(#1)}
\newcommand{\eim}[1]{\catname{EM}(\m{#1})}
\def\H{\textbf {\textup{H}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\newcommand{\fuz}[1]{\textbf{\textup{Fuz}}({\textbf {\textup{#1}}})}
%\newcommand{\yo}{\mathcal{Y}}
\renewcommand*{\bibfont}{\footnotesize}

@@

\begin{frame}{DPO rewriting sytems}\justifying

		We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}).

	\smallskip 
	\begin{overprint}
		\onslide<1|handout:1>
		A \emph{rewriting rule} is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		\onslide<2| handout:0>

@@

@@

		Second: remove the image of $m$  building a pushout square. \vspace{-0.075cm}
		\begin{center}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->, dotted](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->, dotted](C)--(G)node[pos=0.5, below]{$l'$};
			\end{tikzpicture}
		\end{center}		
		\onslide<4| handout:0>

		A \emph{rewriting rule} in a category $\X$ is a pair  of arrows $l:K\to L$ and $r:K\to R$ with the same domain. To rewrite an object $G$ according to a rule we proceed in three steps.

		Third: fill the hole so obtained, with $R$ taking a pushout.\vspace{-0.075cm}
		\begin{center}\hspace{0.045cm}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\node(H)at(4,-2){$H$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->](C)--(G)node[pos=0.5, below]{$l'$};
				\draw[->, dotted ](C)--(H)node[pos=0.5, below]{$r'$};
				\draw[->, dotted ](R)--(H)node[pos=0.5, right]{$q$};
			\end{tikzpicture}
		\end{center}		
	\end{overprint}  	
\end{frame}

\begin{frame}{DPO rewriting system}\justifying
The categorical framework is given by \emph{$\mathcal{M}$-adhesive categories}. \pause 
\begin{definition}[\cite{lack2005adhesive,azzi2019essence,ehrig2014adhesive}]
	Let $\mathcal{M}$ be a class of monos in a category $\X$, containing all isomorphisms, stable under pullbacks and composition. $\X$ is \emph{$\mathcal{M}$-adhesive} if, given a cube below with the hooked arrows in $\mathcal{M}$, the bottom face a pushout and the back faces pullbacks, we have:

@@

		A system is \emph{linear} if every rule is made by arrows in $\mathcal{M}$, \emph{left-linear} if the left hand side of every rule is in $\mathcal{M}$.
	\end{definition}

@@

\pause 
\begin{alertblock}{Motivation}
	Our long term aim is to provide a concurrent semantics for left-linear DPO-rewriting systems (see \cite{baldan2017domains}). 
\end{alertblock}

@@

\begin{frame}{An example}
	If we want a (very sweet) coffee we can add first the white sugar and then the brown one. Finally we can stir it.
	\pause 
\begin{figure}
	\input{Figure/coffee-rewriting2.tex}
\end{figure}\pause

@@

@@

		\item
		\label{lem:indep-global-left:2}
		Suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
		is a switching sequence
		$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
		\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
	\end{enumerate}
\end{lemma}
\end{frame}

@@

	\smallskip \pause

	In \cite{baldan2017domains} a category of domains, called \emph{weak prime domains} has been proposed as a possible concurrent semantic for systems allowing fusions.

	\smallskip \pause We plan to build on the results proved in this paper to show that to every left-linear DPO rewriting system it is possible to associate one of these domains, representing ``derivation up to switches''.
\end{frame}

>>>>

\newcommand{\modd}[0]{\catname{Mod}}
\newcommand{\des}[1]{Des({#1})}
\newcommand{\prof}[2]{\mathsf{Proof}({#1},{#2})}
\newcommand{\ab}[0]{=-Adj}
\newcommand{\ba}[0]{=-E}
\newcommand{\su}[2]{\mathfrak{s}_{#1}^{#2}}
\newcommand{\ipos}[1]{\catname{Pos}(\catname{#1})}
\newcommand{\Su}[2]{\mathfrak{S}_{#1}^{#2}}
\newcommand{\Pro}[2]{\mathfrak{P}_{#1}^{#2}}
\newcommand{\surr}[0]{\mathsf{surr}}
\newcommand{\propag}[0]{\mathsf{propag}}
\newcommand{\mdl}[1]{\catname{Mod}(#1)}
\newcommand{\alg}[1]{\catname{FAlg}(#1)}
\newcommand{\eim}[1]{\catname{EM}(\m{#1})}
\def\H{\textbf {\textup{H}}}
\def\X{\textbf {\textup{X}}}
\def\Y{\textbf {\textup{Y}}}
\newcommand{\fuz}[1]{\textbf{\textup{Fuz}}({\textbf {\textup{#1}}})}
%\newcommand{\yo}{\mathcal{Y}}
%\renewcommand*{\bibfont}{\footnotesize}

@@

\begin{frame}{DPO rewriting sytems}\justifying
	\begin{overprint}
		\onslide<1|handout:1>

		We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}). 
		\smallskip

		\onslide<2|handout:0>

		We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}). 
		\smallskip

		We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}). 
		\smallskip

@@

		\onslide<4| handout:0>

We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}). 
\smallskip

@@

		Second: ``remove'' the image of $m$  building a pushout square. \vspace{-0.075cm}
		\begin{center}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->, dotted](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->, dotted](C)--(G)node[pos=0.5, below]{$l'$};
			\end{tikzpicture}
		\end{center}		
		\onslide<5| handout:0>

We will focus on the \emph{double pushout} (DPO) approach  to  graph rewriting (cfr. \cite{ehrig2006fundamentals}). 
\smallskip

		Third: glue $R$ into the ``hole'',  taking a pushout.\vspace{-0.075cm}
		\begin{center}\hspace{0.08cm}
			\begin{tikzpicture}
				\node(L)at(0,0){$L$};
				\node(K)at(2,0){$K$};
				\node(R)at(4,0){$R$};
				\node(G)at(0,-2){$G$};
				\node(C)at(2,-2){$C$};
				\node(H)at(4,-2){$H$};
				\draw[->](K)--(L)node[pos=0.5, above]{$l$};
				\draw[->](K)--(R)node[pos=0.5, above]{$r$};
				\draw[->](L)--(G)node[pos=0.5, left]{$m$};
				\draw[->](K)--(C)node[pos=0.5, right]{$p$};
				\draw[->](C)--(G)node[pos=0.5, below]{$l'$};
				\draw[->, dotted ](C)--(H)node[pos=0.5, below]{$r'$};
				\draw[->, dotted ](R)--(H)node[pos=0.5, right]{$q$};
			\end{tikzpicture}
		\end{center}		
	\end{overprint}  	
\end{frame}

\begin{frame}{DPO rewriting systems}\justifying
The categorical framework is given by \emph{$\mathcal{M}$-adhesive categories}. \pause 
\begin{definition}[\cite{lack2005adhesive,azzi2019essence,ehrig2014adhesive}]
	Let $\mathcal{M}$ be a class of monos in a category $\X$, containing all isomorphisms, stable under pullbacks and composition. $\X$ is \emph{$\mathcal{M}$-adhesive} if, given a cube below with the hooked arrows in $\mathcal{M}$, the bottom face a pushout and the back faces pullbacks, we have:

@@

		A system is \emph{linear} if every rule is made by arrows in $\mathcal{M}$, \emph{left-linear} if the left leg of every rule is in $\mathcal{M}$.
	\end{definition}

@@

\pause 
\begin{alertblock}{Motivation}
	Our long term aim is to provide a concurrent semantics for left-linear DPO-rewriting systems (see \cite{baldan2021concurrent}). 
\end{alertblock}

@@

\begin{frame}{An example}\justifying
	If we want a (very sweet) coffee we can add first the white sugar and then the brown one. Finally we can stir it.
	\pause 
\begin{figure}
	\input{Figure/coffee-rewriting2.tex}
\end{figure}\pause

@@

\end{frame}
\iffalse 
\begin{frame}{Results: the Three Steps Lemma}\justifying

@@

		\item
		\label{lem:indep-global-left:2}
		Suppose that $\dder{D}_1$ and $\dder{D}_2$ are independent, and that, in $\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_0 \cdot \dder{D}_2' \cdot \dder{D}_1'$, $\dder{D}_0$ and $\dder{D}_2'$ are independent.  If there
		is a switching sequence
		$\dder{D}_0\cdot \dder{D}_1 \cdot \dder{D}_2 \shift{(0,1)}
		\dder{D}_1' \cdot \dder{D}_0' \cdot \dder{D}_2 \shift{(1,2)}
		\dder{D}_1' \cdot \dder{D}_2' \cdot \dder{D}_0''$ then in the last derivation $\dder{D}_1'$ and $\dder{D}_2'$ are independent.
	\end{enumerate}
\end{lemma}
\end{frame}
\fi

@@

	\medskip \pause

	In \cite{baldan2021concurrent} a category of domains, called \emph{weak prime domains} has been proposed as a possible concurrent semantic for systems allowing fusions.

	\medskip \pause We plan to build on the results proved in this paper to show that to every left-linear DPO rewriting system it is possible to associate one of these domains, representing ``derivation up to switches''.
\end{frame}
